Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-03-12 draft.tex
made nice pull plots
fuckll
tiny typo fix
tiny typo fix 2
proof read up to summary

\begin{figure}[b]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S5_200.png}
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S7_200.png}
        \caption{Pull distribution for the angular observables $S_5$ (upper) and $S_7$ (lower), extracted from $2\cdot 10^5$ studies of
        $200$ simulated events each for the decay $B\to K^*\ell^+\ell^-$. The red curve represents a fit to a Gaussian distribution.}
        \label{fig:pulls}
\end{figure}

@@

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/S5_scat.pdf}
            \includegraphics[width=0.45\textwidth]{figs/S7_scat.pdf}
            \caption{The joint distribution of the two estimators
                $\est{S}_i^\text{MoM}$ and $\est{S}_i^\text{Fit}$ that arise
                from the method of moments and a maximum-likelihood fit,
                respectively. We show the estimators for both of our benchmark
                observables $S_3$ (upper) and $S_5$ (lower).
                \danny{Marcin: please replace by $S_5$ (upper) and $S_7$ (lower).}
            }
        \label{fig:correlation}
\end{figure}

@@

We have carried out a combined analytical and numerical study of the method of moments; a method for
the extraction of angular observales from the angular distribution of a general multi-body process.
We have studied the performance of the method of moments using pseudo data derived from the SM
predctions for one uniangular decay ($B\to K\ell^+\ell^-$) and one triangular decay ($B\to K^*\ell^+\ell^-$).
From this, we find rapid convergence of the joint likelihood of the angular observables towards a multivariate Gaussian.
We draw the conclusion that this method exhibits several benefits in the determination of angular observables when
compared with a maximum-likelihood fit.\\

First, we find no bias in the determinations of the angular observables even for a small number of events.
However, this comes at the expense of a slight increase of the order of $10\%$ for the statistical uncertainty
on the mean.
Second, in the absence of detector effects, a determination using the method of moments is correct and unbiased
even when not all aspects of the physical distribution are modeled. This is explicitly shown for the case of
higher partial waves in multibody final states.
Third, we develop a systematic method for the dertermination of detector effects that lead to
dilution and mixing of the angular observables. We present an algorithm to calculate the necessary
unfolding matrix, which is only computationally feasible when using the method of moments.
Fourth, the fact that the joint distribution of the angular observables that results from the method of moments
is a multivariate Gaussian distribution allows for an easy and precise transfer of correlation
information to subsequent theoretical analyses.
Last but not least, the resulting distribution arises without the need for additional constraints
in order to achieve convergence, such as e.g. in some maximum-likelihood fits. Therefore, results obtained
using the method of moments can be directly averaged, without the need to remove
the constraints a posteriori.\\

In conclusion, we argue that the method of moments is a competetive alternative to
maximum-likelihood fits as long as angular distributions are involved. We wish to raise
the interesting prospect of extending this method to applications that feature
PDFs composed from non-angular orthogonal bases.

>>>>

\begin{figure}[b]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/Q2_5_6_S5_200.pdf}
            \includegraphics[width=0.45\textwidth]{figs/Q2_5_6_S9_200.pdf}
        \caption{Pull distribution for the angular observables $S_5$ (upper) and $S_7$ (lower), extracted from $2\cdot 10^5$ studies of
        $200$ simulated events each for the decay $B\to K^*\ell^+\ell^-$. The red curve represents a fit to a Gaussian distribution.}
        \label{fig:pulls}
\end{figure}

@@

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/S5_scat.pdf}
            \includegraphics[width=0.45\textwidth]{figs/S9_scat.pdf}
            \caption{The joint distribution of the two estimators
                $\est{S}_i^\text{MoM}$ and $\est{S}_i^\text{Fit}$ that arise
                from the method of moments and a maximum-likelihood fit,
                respectively. We show the estimators for both of our benchmark
                observables $S_3$ (upper) and $S_5$ (lower).
            }
        \label{fig:correlation}
\end{figure}

@@

We have carried out a combined analytical and numerical study of the method of
moments; a method for the extraction of angular observables from the angular
distribution of a general multi-body process.  We have studied the performance
of the method of moments using pseudo data derived from the SM predictions for
one uniangular decay ($B\to K\ell^+\ell^-$) and one triangular decay ($B\to
K^*\ell^+\ell^-$).  From this, we find rapid convergence of the joint
likelihood of the angular observables towards a multivariate Gaussian.  We draw
the conclusion that this method exhibits several benefits in the determination
of angular observables when compared with a maximum-likelihood fit.\\

First, we find no bias in the determinations of the angular
observables even for a small number of events.  However, due to fewer
model assumptions, the uncertainty on the mean values increases by
roughly $10\%$--$30\%$ compared
to likelihood fits.\\
Second, in the absence of detector effects, the method of moments does
not rely on model assumptions for the partial-wave composition of the
PDF. This is explicitly shown for the case of higher
partial waves in multibody final states.\\
Third, we develop a systematic method for the determination of
detector effects that lead to dilution and mixing of the angular
observables. We present an algorithm to calculate the necessary
unfolding matrix, which is computationally feasible only when using
the method
of moments. The algorithm also accounts for higher partial waves.\\
Fourth, the joint distribution of the angular observables resulting
from the method of moments is well approximated by a multivariate
Gaussian distribution even for small number of events $N \sim 30$ both
for the ideal uniform acceptance and a realistic example.  This
facilitates the precise transfer of correlation information
to subsequent theoretical analyses.\\
Last but not least, the resulting distribution arises without the need
for additional model constraints. Thus more observables can be
inferred from the same data than in a likelihood fit. In addition,
the results from the method of moments can be more easily averaged or combined; e.g., in global fits.\\

In conclusion, we argue that the method of moments is a competitive alternative to
maximum-likelihood fits if angular distributions are involved. We wish to raise
the interesting prospect of extending this method to applications that feature
PDFs composed from non-angular orthogonal bases.
